---
title: "# Week5-Lab-Assignment"
subtitle: "Chronic Kidney Disease Data Analysis"
author: "Kasra Vand"
date: "2024-09-27"
output: pdf_document

fontsize: 11pt
geometry: margin=1in

header-includes:
  - \definecolor{NavyBlue}{RGB}{0,0,128}
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Loading and Preparation

First, let's load the dataset and prepare it for analysis.

```{r load_data}
# Load required libraries
library(dplyr)

# Read the dataset
ckd_data <- read.csv("Documents/Chronic_Kidney_Dsease_data.csv")

# Display the first few rows and summary
# head(ckd_data)
# summary(ckd_data)
```

## Binomial Distribution

Let's look at the prevalence of certain risk factors in our dataset.

1. Family History of Kidney Disease

```{r family_history}
prob_family_history <- mean(ckd_data$FamilyHistoryKidneyDisease)

# Probability of exactly 10 out of 50 patients having family history
dbinom(10, 50, prob_family_history)

# Probability of at most 10 out of 50 patients having family history
pbinom(10, 50, prob_family_history)

# Probability of more than 10 out of 50 patients having family history
pbinom(10, 50, prob_family_history, lower.tail = FALSE)
```

Questions:
a) What is the probability that exactly 10 out of 50 patients have a family history of kidney disease?
b) What is the probability that at most 10 out of 50 patients have a family history of kidney disease?
c) What is the probability that more than 10 out of 50 patients have a family history of kidney disease?

Answers:
a) The probability that exactly 10 out of 50 patients have a family history of kidney disease is approximately 0.0731 or 7.31%.
b) The probability that at most 10 out of 50 patients have a family history of kidney disease is approximately 0.9141 or 91.41%.
c) The probability that more than 10 out of 50 patients have a family history of kidney disease is approximately 0.0859 or 8.59%.

2. Simulation: Smoking Status

Let's simulate sampling 100 patients and count how many are smokers.

```{r smoking_simulation}
prob_smoking <- mean(ckd_data$Smoking)
num_patients <- 100
num_replicates <- 1000

set.seed(2023)

simulate_smoking <- function() {
  sum(sample(c(0,1), size = num_patients, 
             prob = c(1 - prob_smoking, prob_smoking), 
             replace = TRUE))
}

smoking_counts <- replicate(num_replicates, simulate_smoking())

hist(smoking_counts, breaks = 20, main = "Distribution of Smokers in 100 Patient Samples")
```

Question:
d) Based on the simulation results, describe the distribution of smokers in 100 patient samples.

Answer:
d) The histogram shows the distribution of smokers in 100 patient samples over 1000 simulations. The distribution appears to be roughly symmetric and bell-shaped, centered around the mean number of smokers. This suggests that the number of smokers in a sample of 100 patients follows an approximately normal distribution.

## Normal Distribution

3. Analysis of GFR (Glomerular Filtration Rate)

```{r gfr_analysis}
mean_gfr <- mean(ckd_data$GFR)
sd_gfr <- sd(ckd_data$GFR)

# Probability of a patient having GFR less than 60 (indicating kidney disease)
pnorm(60, mean_gfr, sd_gfr)

# 95th percentile of GFR in the population
qnorm(0.95, mean_gfr, sd_gfr)
```

Questions:
e) What is the probability of a patient having a GFR less than 60 (indicating kidney disease)?
f) What is the 95th percentile of GFR in the population?

Answers:
e) The probability of a patient having a GFR less than 60 (indicating kidney disease) is approximately 0.4101 or 41.01%.
f) The 95th percentile of GFR in the population is approximately 116.26 mL/min/1.73m².

4. Blood Pressure Analysis

```{r bp_analysis}
# Assuming hypertension is defined as SystolicBP >= 140 or DiastolicBP >= 90
hypertension <- (ckd_data$SystolicBP >= 140) | (ckd_data$DiastolicBP >= 90)
prob_hypertension <- mean(hypertension)

# Probability of at least 40 out of 100 patients having hypertension
pbinom(39, 100, prob_hypertension, lower.tail = FALSE)
```

Question:
g) What is the probability of at least 40 out of 100 patients having hypertension?

Answer:
g) The probability of at least 40 out of 100 patients having hypertension is 1, or 100%. This suggests that hypertension is very common in this patient population, with the prevalence being much higher than 40%.

## Poisson Distribution

5. Urinary Tract Infections

Assuming UTIs are rare events, we can model them using a Poisson distribution.

```{r uti_analysis}
mean_uti <- mean(ckd_data$UrinaryTractInfections)

# Probability of a patient having more than 2 UTIs in a year
ppois(2, mean_uti, lower.tail = FALSE)

# Expected number of UTIs in a group of 100 patients
lambda_100 <- mean_uti * 100
lambda_100

# Probability of more than 50 UTIs in a group of 100 patients
ppois(50, lambda_100, lower.tail = FALSE)
```

Questions:
h) What is the probability of a patient having more than 2 UTIs in a year?
i) What is the expected number of UTIs in a group of 100 patients?
j) What is the probability of more than 50 UTIs occurring in a group of 100 patients?

Answers:
h) The probability of a patient having more than 2 UTIs in a year is approximately 0.00133 or 0.133%.
i) The expected number of UTIs in a group of 100 patients is approximately 21.04.
j) The probability of more than 50 UTIs occurring in a group of 100 patients is extremely low, approximately 2.33e-08 or 0.00000233%.
```
